---
title: "Exercise 7: Predicting Ames house prices with tidymodels"
date: "`r Sys.Date()`"
author: "Ranji Raj"
output: 
  html_document:
    fig_width: 6
    fig_height: 5
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
library(tidyverse)
library(tidymodels)
library(AmesHousing)

(ames <- make_ames() %>% select(-matches("Qu")))
```

## Task 1 - parsnip

1. Create a simple linear regression model on the Ames data to predict _sale price_ (`Sale_Price`) based on _above ground living area_ (`Gr_Liv_Area`).
2. Apply the model to the original data.
3. Use `mutate()` to add a column with the observed sale prices; name it `truth`.

```{r t1, eval = FALSE}
lm_spec <- linear_reg() %>% 
  set_engine(engine = "lm") %>%
  set_mode(mode = "regression")

lm_fit <- fit(object = lm_spec,
              formula = Sale_Price ~ Gr_Liv_Area, 
              data = ames)

price_pred <- lm_fit %>% 
  predict(new_data = ames) %>% 
  mutate(truth = ames$Sale_Price)

price_pred
```

## Task 2 - holdout validation

Use `initial_split()`, `training()`, `testing()`, `lm()` and `rmse()` to:

1. Split `ames` into a training (70%) and a test (30%) set. Save the `rsplit`.
2. Extract the training data. Fit a linear model to it. Save the model as `lm_fit`.
3. Measure the RMSE of your linear model with your test set.

Keep `set.seed(100)` at the start of your code.

```{r t2, eval = FALSE}
set.seed(100)

ames_split <- initial_split(ames, prop = 0.7)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

lm_fit <- fit(lm_spec, Sale_Price ~ Gr_Liv_Area, data = ames)

price_pred  <- lm_fit %>% 
  predict(new_data = ames_test) %>% 
  mutate(truth = ames_test$Sale_Price)

rmse(price_pred, truth = truth, estimate = .pred)
```

## Task 3: k-fold cross validation

Modify the code below to return the **Mean Absolute Error** and **R²**.
Visit <https://yardstick.tidymodels.org/reference/index.html#section-regression-metrics> to find the right function to use.

```{r t3, eval = FALSE}
set.seed(100)

folds <- vfold_cv(ames, v = 5)

fit_res <- fit_resamples(object = lm_spec,
                         preprocessor = Sale_Price ~ Gr_Liv_Area,
                         resamples = folds,
                         control = control_resamples())

collect_metrics(fit_res, summarize = TRUE)
```

## Task 4: bootstrap validation - decision tree vs knn

Compare the performance of a regression tree model with a k-nearest neighbor
model on 10 bootstrap samples using RMSE as evaluation measure.

Set `rpart` as engine for the decision tree model and `kknn` as engine for the
k-nearest neighbor model. Leave all model-specific parameters (e.g. `tree_depth`
and `neighbors`) at default.

Which of the two methods perform better and why?

```{r t4, eval = FALSE}
set.seed(100)
bs <- vfold_cv(ames, times = 10)

#decision tree model
rt_spec <- decision_tree() %>%          
  set_engine(engine = "rpart") %>% 
  set_mode("regression")

fbs <- fit_resamples(rt_spec, Sale_Price ~ Gr_Liv_Area, resamples = bs)
fbs %>% collect_metrics()

#k-nearest neighbor model
knn_spec <- nearest_neighbor() %>%          
  set_engine(engine = "kknn") %>% 
  set_mode("regression")

fknn <- fit_resamples(knn_spec, Sale_Price ~ Gr_Liv_Area, resamples = bs)
fknn %>% collect_metrics()
```
`regression tree` model performs better. The RMSE for `std_err` is comparatively lower.


## Task 5: investigating non-default hyperparameter values

Create a new classification tree model specification; name it `big_tree_spec`.
Set the cost complexity to 0, and the minimum number of data points in a node to split to 1.

Compare the metrics of the big tree to the vanilla tree.
Which one predicts the test set better?

Vanilla tree performance:

- accuracy = 0.64
- ROC AUC = 0.66

```{r t5, eval = FALSE}
stackoverflow <- read_rds(here::here("C:/Users/User/Downloads/datascir21/ex/ex-7/stackoverflow.rds"))
set.seed(100)
so_cv <- vfold_cv(stackoverflow, v = 5)

big_tree_spec <- decision_tree(min_n = 1, cost_complexity = 0) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

fit_resamples(big_tree_spec,
              remote ~ .,
              resamples = so_cv) %>%
  collect_metrics()
```
Better predictor on the test set: Vanilla tree as the ROC AUC value is better.


## Task 6: hyperparameter tuning

Create a KNN workflow on the Ames house prices problem that tunes over the number of neighbors in a 5-fold stratified CV scheme.
The candidate values of `neighbors` are `1, 11, 21, 31, ..., 101`.

```{r t6, eval = FALSE}
k_grid <- expand_grid(neighbors = seq(1,101,10))

knn_tuner <- nearest_neighbor(neighbors = tune()) %>% 
  set_engine("kknn") %>% 
  set_mode("regression")

knn_twf <- workflow() %>% 
  add_model(knn_tuner) %>%
  add_formula(Sale_Price ~ Gr_Liv_Area)

set.seed(100)
cv_folds <- vfold_cv(ames, v = 5, strata = Sale_Price, breaks = 4)
knn_results <- knn_twf %>% tune_grid(resamples = cv_folds, grid = k_grid) 

knn_rmse <- knn_results %>% 
  collect_metrics() %>% 
  filter(.metric == "rmse")
knn_rmse
```

## Task 7: z-scoring with recipes

Create a recipe that performs z-score normalization on each numeric variable of the ames data.

A z-score is calculated by subtracting the variable mean from an individual raw value and then dividing the difference by the variable standard deviation.

Tip: Look up the appropriate `step_*()` functions at <https://tidymodels.github.io/recipes/reference/index.html>.

```{r t7, eval=FALSE}
recipe(Sale_Price ~ ., data = ames) %>% 
  step_normalize(all_numeric())
```

## Task 8: preprocessing pipelines

Write a recipe for the Ames data that:

1. adds a novel level to all factors
2. converts all factors to dummy variables
3. catches any zero variance variables
4. centers all of the predictors
5. scales all of the predictors
6. computes the first 3 principal components

Save the result as `pca_rec`.

```{r t8, eval=FALSE}
pca_rec <- 
  recipe(Sale_Price ~ ., data = ames) %>%
  step_novel(all_nominal()) %>%
  step_dummy(all_nominal()) %>%
  step_zv(all_predictors()) %>%
  step_center(all_predictors()) %>%
  step_scale(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 3)
pca_rec
```

