---
title: "The R language: Vectors, Classes, Functions, Iteration"
date: "`r Sys.Date()`"
author: "Ranji Raj"
output: 
  html_document:
    fig_width: 6
    fig_height: 5
    toc: true
    toc_float: true
---

```{r setup}
library(tidyverse)
library(lubridate)

data(attrition, package = "modeldata")
attrition <- as_tibble(attrition)
```

## Functions

Write your own standard deviation function without using `mean()`, `var()` and `sd()`.

```{r}
#Mean
Ranji.Mean <- function(x) {
  res <- sum(x) / length(x)
  return(res) 
  }

Ranji.Mean(0:10)

#Variance
Ranji.Var = function(x) {
  m <- sum(x) / length(x)
  return(
    sum((x-m)^2)/(length(x) - 1)
    )
}

Ranji.Var(0:10)

#S.D.
Ranji.Sd = function(x) {
    sqrt(sum((x - (sum(x) / length(x)))^2) / (length(x)-1))
}
Ranji.Sd(0:10)
```


Write a greeting function that takes two arguments, a time and a name, and says "good morning <name>", "good afternoon <name>", or "good evening <name>", depending on the time of day. 
Use `lubridate::now()` as default for the time argument.

```{r}
greet <- function(time = lubridate::now(), name) {
  hour <- parse_number(str_sub(time, 12, 13))
  if(hour < 0 || hour > 23) {
    stop("Please define a correct time")
  }
  if (hour < 12) {
    cat("good morning", name)
  } else if (hour < 17) {
    cat("good afternoon", name)
  } else {
    cat("good evening", name)
  }
}

greet(name = "Ranji")
#> [1] "good afternoon Ranji"
greet(ymd_h("2021-03-31 05"), "Bob")
#> [1] "good morning Bob"
greet(ymd_h("2021-03-31 13"), "Alice")
#> [1] "good afternoon Alice"
greet(ymd_h("2021-03-31 20"), "Cedric")
#> [1] "good evening Cedric"
```


## Iteration

Use a map function to check for each column of `attrition` whether it is a factor.

```{r attrition-is-factor}

attrition %>% map_df(~(data.frame(class = class(.x))),
                     .id = "Column") %>%
  mutate(is_factor = class == 'factor') %>% 
  select(Column, is_factor)
```

Use a map function to extract the type of each column in `attrition`.

```{r}
attrition %>% map_df(~(data.frame(class = class(.x))),
                     .id = "Column")
```

Using a map function, write a pipeline which returns the number of distinct values for each factor column in `attrition`. (Hint: you can select columns based on a logical condition using `keep()`).

```{r}
attrition %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "Column") %>% 
  mutate(is_factor = class == 'factor') %>%
  filter(is_factor == TRUE) %>% 
  select(Column, n_distinct)

```

Simplify the following code using `map_dfr()`.

```{r}
setosa <- read_csv("iris-setosa.csv") %>%
  mutate(Species = "setosa")
versicolor <- read_csv("iris-versicolor.csv") %>%
  mutate(Species = "versicolor")
virginica <- read_csv("iris-virginica.csv") %>%
  mutate(Species = "virginica")
df_iris <- bind_rows(setosa, versicolor, virginica)
df_iris
```

```{r}
species <- c("setosa", "versicolor", "virginica")
map_dfr(species, get)
```

Complete the following pipeline to create an age histogram for each Department. 
Set the number of histogram bins to 12.

```{r}
attrition %>%
  nest(data = -Department) %>%
  mutate(gg = map(data, ~ ggplot(.x) + aes(Age) + geom_histogram(bins = 12))) %>%
  pull(gg)
```

The code below builds a linear regression model for the relationship between age and monthly income for each department in the attrition data. 
Use a map function to add a column `r_squared` by extracting the real-valued `r.squared` element from `lm_summary` which quantifies goodness of the fit. (See `?lm.summary`)

```{r}
attrition %>%
  group_by(Department) %>%
  summarise(lm_summary = list(summary(lm(MonthlyIncome ~ Age)))) %>%
  mutate(r_squared = map_dbl(lm_summary, "r.squared"))
```
_Other alternatives_

```{r}
attrition %>%
    group_by(Department) %>%
    summarise(lm_summary = list(summary(lm(MonthlyIncome ~ Age)))) %>%
    mutate(r_squared = purrr::map_dbl(lm_summary, function(x) x[["r.squared"]]))
```

```{r}
attrition %>%
  group_by(Department) %>%
  summarise(lm_summary = list(summary(lm(MonthlyIncome ~ Age))), 
            r_squared = map_dbl(lm_summary, pluck, "r.squared"))
```
